\thetitle: Technical Report

### 2023-04-05 metatheory2.tex
id
Update Acknowledgement
Removing mentioning of the appendix.

Unlike in \cref{sec:metatheory2}, which uses named binders, in this appendix we
use \citeauthor{debruijn1964} indices and \emph{explicit
substitutions}~\cite{abadi1989subst,substcalculus}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}, which
applies~$\gm$ to the type~$T$) are part of the syntax rather than
meta-operations.
We work in an intrinsically typed setting: terms are well typed by construction.
%
Consequently, we omit without ambiguity some obvious premises needed for well-formedness.
% 
This style of syntax formulation follows a recent
trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
known as the ``algebraic presentation'' of MLTT.
%We introduce judgmental equality ($\equiv$)
%for derivations (of all kinds of judgements). In other words, we will have judgemental equality between terms, 
%contexts, explicits substitutions, and etc. 
%This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT. 
Moreover, universe levels are explicit in this appendix. Other aspects of the syntax remain the same as in \cref{sec:metatheory2}.

@@

% Furthermore, the semantic model we develop in this appendix respects these
% judgmental equalities by construction.

@@

\Citeauthor{debruijn1964} indices and explicit substitutions make details about
binders and substitutions explicit.
%
Using explicit substitutions obviates the need for special treatment of
substitutions in the proofs, as substitutions are part of the syntax.
%The change of the syntax formulation is important to the simplicity of our
%proof, as we want to handle the proof about substitutions and terms
%uniformly---thus making substitution as part of the syntax (explicit
%substitution) will be helpful. 
%
The form \var{n} represents a variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
%
Substitutions are typed with the form 
$\goodSub{\Gm}{\gm}{\Dl}$.
The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
%
The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
where $\SubstId$ is the identity substitution (\ruleref{sub/id}).

>>>>

Unlike in \cref{sec:metatheory2}, which uses named binders, here we
use \citeauthor{debruijn1964} indices and \emph{explicit
substitutions}~\cite{abadi1989subst,substcalculus}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}, which
applies~$\gm$ to the type~$T$) are part of the syntax rather than
meta-operations.
We work in an intrinsically typed setting: terms are well typed by construction.
%
Consequently, we omit without ambiguity some obvious premises needed for well-formedness.
% 
This style of syntax formulation follows a recent
trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
known as the ``algebraic presentation'' of MLTT.
%We introduce judgmental equality ($\equiv$)
%for derivations (of all kinds of judgements). In other words, we will have judgemental equality between terms, 
%contexts, explicits substitutions, and etc. 
%This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT. 
Moreover, universe levels are explicit here. Other aspects of the syntax remain the same as in \cref{sec:metatheory2}.

@@

% Furthermore, the semantic model we develop here respects these
% judgmental equalities by construction.

@@

\Citeauthor{debruijn1964} indices and explicit substitutions make details about
binders and substitutions explicit.
%
Using explicit substitutions obviates the need for special treatment of
substitutions in the proofs, as substitutions are part of the syntax.
%The change of the syntax formulation is important to the simplicity of our
%proof, as we want to handle the proof about substitutions and terms
%uniformly---thus making substitution as part of the syntax (explicit
%substitution) will be helpful. 
%
The form \var{n} represents a variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
%
Substitutions are typed with the form 
$\goodSub{\Gm}{\gm}{\Dl}$.
The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
%
The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
where $\SubstId$ is the identity substitution (\ruleref{sub/id}).
We occasionally use the notation "id" for $\SubstId$.

### 2023-04-05 stlc-linkTran-typing.tex
Fix Wsup - Wsup needs specify the index of the constructor.
Removing mentioning of the appendix.

In this appendix, we sketch how to use the ``library'' of linkage
transformers in \TT to inductively construct a linkage transformer that models a
derived family of \lsti{STLC}.

@@

% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{centered}
\small
\renewcommand*{\arraystretch}{1.25}
\begin{tabular}{|l|l|l|}
\hline
\rowcolor[HTML]{FFFC9E} 
surface-syntax program & $i$      & $\goodInh{\cdot}{h_i}{\lsig_i}{\lsig_i'}$             \\ \hline
\lsti!Family STLCBool extends STLC.!   & 0        & $\InhId$                                             \\ \cline{2-3} 
\lsti!FInductive tm += !       & 1        & $\InhOv{h_0}{\wcode{\wsig_{\texttt{tm}}'}}$                      \\ \cline{2-3} 
\codecomment{existing constructors}   & 2        & $\InhOv{h_1}{\Wsup{\wsig_{\texttt{tm}}'}{\top}{\bot}}$           \\ \cline{2-3} 
\lsti!| tm_true |!\ \;\dadada    & 3        & $\InhExt{h_2}{\Wsup{\wsig_{\texttt{tm}}'}{\top}{\bot}}$          \\ \cline{2-3} 
\rowcolor[HTML]{CDCDCD} 
\lsti!FRecursion subst!\ \;\dadada\ \;\lsti!+=! & $\alpha$ & $\InhId$ \\ \cline{2-3} 
\rowcolor[HTML]{CDCDCD} 
\lsti!  Case tm_true := !\dadada. \lsti! Case !\dadada & $\beta$  & $\InhExt{h_\alpha}{...}$ \\ \cline{2-3} 
\rowcolor[HTML]{FFFFFF} 
\lsti!End subst.!              & 4        & $\InhNest{h_3}{h_\beta}$
  %with $\uparrow^s_3$
  \\ \cline{2-3} 
                          & 5        & $\InhOv{h_4}{\lambda t. \Wrec{\wsig_{\texttt{tm}}'}{t}{"substCases"}}$ \\ \cline{2-3} 
\lsti!FInductive ty += !\dadada. & 6        & ...                                                  \\ \cline{2-3} 
\codecomment{Inherit "env"}   & 7        & $\InhInh{h_6}$
  %with $\uparrow^s_6$
  \\ \cline{2-3} 
\codecomment{Inherit "empty"}   & 8        & $\InhInh{h_7}$
  \\ \cline{2-3} 
\dadada   & \dadada & \dadada
  \\ \hline
\end{tabular}
\end{centered}

>>>>

Here, we sketch how to use the ``library'' of linkage
transformers in \TT to inductively construct a linkage transformer that models a
derived family of \lsti{STLC}.

@@

% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{centered}
\small
\renewcommand*{\arraystretch}{1.25}
\begin{tabular}{|l|l|l|}
\hline
\rowcolor[HTML]{FFFC9E} 
surface-syntax program & $i$      & $\goodInh{\cdot}{h_i}{\lsig_i}{\lsig_i'}$             \\ \hline
\lsti!Family STLCBool extends STLC.!   & 0        & $\InhId$                                             \\ \cline{2-3} 
\lsti!FInductive tm += !       & 1        & $\InhOv{h_0}{\wcode{\wsig_{\texttt{tm}}'}}$                      \\ \cline{2-3} 
\codecomment{existing constructors}   & 2        & $\InhOv{h_1}{\Wsup{}{\wsig_{\texttt{tm}}'}{\top}{\bot}}$           \\ \cline{2-3} 
\lsti!| tm_true |!\ \;\dadada    & 3        & $\InhExt{h_2}{\Wsup{}{\wsig_{\texttt{tm}}'}{\top}{\bot}}$          \\ \cline{2-3} 
\rowcolor[HTML]{CDCDCD} 
\lsti!FRecursion subst!\ \;\dadada\ \;\lsti!+=! & $\alpha$ & $\InhId$ \\ \cline{2-3} 
\rowcolor[HTML]{CDCDCD} 
\lsti!  Case tm_true := !\dadada. \lsti! Case !\dadada & $\beta$  & $\InhExt{h_\alpha}{...}$ \\ \cline{2-3} 
\rowcolor[HTML]{FFFFFF} 
\lsti!End subst.!              & 4        & $\InhNest{h_3}{h_\beta}$
  %with $\uparrow^s_3$
  \\ \cline{2-3} 
                          & 5        & $\InhOv{h_4}{\lambda t. \Wrec{\wsig_{\texttt{tm}}'}{t}{"substCases"}}$ \\ \cline{2-3} 
\lsti!FInductive ty += !\dadada. & 6        & ...                                                  \\ \cline{2-3} 
\codecomment{Inherit "env"}   & 7        & $\InhInh{h_6}$
  %with $\uparrow^s_6$
  \\ \cline{2-3} 
\codecomment{Inherit "empty"}   & 8        & $\InhInh{h_7}$
  \\ \cline{2-3} 
\dadada   & \dadada & \dadada
  \\ \hline
\end{tabular}
\end{centered}
